### 2018-04-13 paper/appendix.tex
pass over all except proof of correctness

\section{Generating a Minimal CDD Expansion}
\label{app:cons-cdd}
% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A recruions-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A minimal CDD expansion $S'$ of $S$ and
    correspondence from $S'$ to $S$.} %
  \myproc{$\expand(S)$ %
    \label{line:expand-begin} }{ %
    % auxiliary procedure:
    \myproc{$\expandaux(S')$
      \label{line:expand-aux-begin} }{ %
      % case split: optional sharing clause:
      \Switch{$\sharingclause(S')$}{ %
        % case: no siblings with overlapping descendants
        \lCase{$\none$:}{ \Return{$S'$} %
          \label{line:expand-ret} } %
        % case: siblings with 
        \lCase{$C \in S', % 
          P \in \predof{S'}$: }{ %
          % expand the return:
          \Return{$\expandaux( %
            \copyrel(S', C, P))$ } %
          \label{line:expand-recurse}
        } %
      } %
    \label{line:expand-aux-end} } %
    % base call: 
    \Return{$( \expandaux(S), \corr )$ %
      \label{line:expand-base-call}}
  } %
  %
  \caption{$\expand$:
    given a recursion-free CHC system $S$, returns a minimal
    CDD expansion $S'$ of $S$ and its correspondence.}
  \label{alg:expand}
\end{algorithm}
% walk through top-level procedure:
\autoref{alg:expand}, given a recruion-free CHC system $S$
(\autoref{line:expand-begin}), returns an minimal CDD expansion $S'$ of
$S$ with a correspondence from $S'$ to
$S$.
%
$\expand$ defines a procedure $\expandaux$
(\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
takes a CHC system $S$ and returns an minimal CDD expansion of 
$S$.
%
$\expand$ runs $\expandaux$ on $S$ and
returns the result, paired with the map $\corr: \predof{S'} \to \predof{S}$ 
(\autoref{line:expand-base-call}).

% walk through aux procedure:
$\expandaux$, given a recursion-free CHC system $S'$,
runs a procedure $\sharingclause$ on $S'$, which tries to
find clause $C \in S'$ and predicate $P \in \bodyof{C}$
that $P$ is in both transitive dependences of two predicates $Q_0$ and $Q_1$ are siblings 
in $\predof{S}'$.
%
In such a case, we say that $(C, P)$ is a
\emph{sibling-shared dependence}.

% case: no sibling-shares
If $\sharingclause$ determines that no sibling-shared dependence
exists, then $\expandaux$ returns $S'$
(\autoref{line:expand-ret}).

% case: there is a sibling-shared dependence:
Otherwise, for sibling-shared dependence $(C, P)$ returned
by $\sharingclause$, $\expandaux$ runs $\copyrel $ on
$S'$, $C$, and $P$, which returns an expansion of
$S'$ by creates fresh copy $P'$ of $P$ and updates $\bodyof{C}$ 
to avoid transitive dependents shared by $Q_0$ and $Q_1$.
%
$\expandaux$ recurses on the CHC system returned by $\copyrel $ and returns the result
(\autoref{line:expand-recurse}).
%

@@

\section{Proof of characterization of CDD systems}
\label{app:char}
The following is a proof of \autoref{thm:cdd-contains}.
%
\begin{proof}
  % proof structure:
  To prove that the union of the classes of linear systems and
  body-disjoint systems are strictly contained by the class of CDD
  systems, we prove %
  \textbf{(1)} the class of linear systems are contained by the class
  of CDD systems, %
  \textbf{(2)} the class of body-disjoint systems are contained by the
  class of CDD systems, and %
  \textbf{(3)} there is some CDD system that is not linear or
  body-disjoint.

  % contains linear systems:
  For goal \textbf{(1)}, let $S$ be an arbitrary linear
  system.
  %
  $S$ is CDD if for each clause $C$ in
  $S$, each pair of distinct predicates in
  the body of $C$, the transitive dependence are disjoint, and
  no predicates shows more than once in the body of $C$.
  (\autoref{defn:cdds}).
  %
  Let $C$ be an arbitrary clause in $S$.
  %
  The body of $C$ has no pair of distinct predicates or a predicates shows more than once, by the fact that
  $S$ is linear.
  %
  Therefore, $S$ is CDD.

  % contains body-disjoint systems:
  For goal \textbf{(2)}, let $S$ be an arbitrary
  body-disjoint system.
  %
  The dependence relation of $S$ is a tree $T$, by the
  definition of a body-disjoint system.
  %
  Let $C$ be an arbitrary clause in $S$, with
  distinct relational predicates $R_0$ and $R_1$ in its body.
  %
  All dependences of $R_0$ and $R_1$ are in subtrees of $T$, which
  are disjoint by the definition of a tree.
  %
  Thus, $S$ is CDD, by \autoref{defn:cdds}.

  % distinguishing example:
  For goal \textbf{(3)}, the system $\mcchc$ is CDD, but is not linear
  or body-disjoint.
\end{proof}

>>>>

\section{Generating a Minimal CDD Expansion}
\label{app:cons-cdd}
% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A recruions-free CHC system $S$.}
  % Output: inductive invariants.
  \Output{A minimal CDD expansion $S'$ of $S$ and
    correspondence from $S'$ to $S$.} %
  \myproc{$\expand(S)$ %
    \label{line:expand-begin} }{ %
    % auxiliary procedure:
    \myproc{$\expandaux(S')$
      \label{line:expand-aux-begin} }{ %
      % case split: optional sharing clause:
      \Switch{$\sharingclause(S')$}{ %
        % case: no siblings with overlapping descendants
        \lCase{$\none$:}{ \Return{$S'$} %
          \label{line:expand-ret} } %
        % case: siblings with 
        \lCase{$C \in S', % 
          P \in \predof{S'}$: }{ %
          % expand the return:
          \Return{$\expandaux( %
            \copyrel(S', C, P))$ } %
          \label{line:expand-recurse}
        } %
      } %
    \label{line:expand-aux-end} } %
    % base call: 
    \Return{$( \expandaux(S), \corr )$ %
      \label{line:expand-base-call}}
  } %
  %
  \caption{$\expand$:
    given a recursion-free CHC system $S$, returns a minimal
    CDD expansion $S'$ of $S$ and its correspondence.}
  \label{alg:expand}
\end{algorithm}
%
% walk through top-level procedure:
Given a recursion-free CHC system $S$, \autoref{alg:expand} returns a
minimal CDD expansion of $S$ (\autoref{defn:min-expansion}).
%
$\expand$ defines a procedure $\expandaux$
(\autoref{line:expand-aux-begin}---\autoref{line:expand-aux-end}) that
takes a CHC system $S$ and returns a minimal CDD expansion of $S$.
%
$\expand$ runs $\expandaux$ on $S$ and
returns the result, paired with the map $\corr: \predof{S'} \to \predof{S}$ 
(\autoref{line:expand-base-call}).

% walk through aux procedure:
$\expandaux$, given a recursion-free CHC system $S'$,
runs a procedure $\sharingclause$ on $S'$, which tries to
find a clause $C \in S'$ and a predicate $P \in \bodyof{C}$
such that $P$ is in the transitive dependencies of two sibling
predicates.
%
In such a case, we say that $(C, P)$ is a
\emph{sibling-shared dependency}.

% case: no sibling-shares
If $\sharingclause$ determines that no sibling-shared dependency
exists, then $\expandaux$ returns $S'$
(\autoref{line:expand-ret}).

% case: there is a sibling-shared dependency:
Otherwise, $\sharingclause$ must have located a sibling-shared
dependency $(C, P)$. In this case, $\expandaux$ runs $\copyrel $ on
$S'$, $C$, and $P$, which returns an expansion of $S'$ by creating a
fresh copy of $P$ and updating $\bodyof{C}$ to avoid the shared
dependency.
%
$\expandaux$ recurses on this expansion and returns the result
(\autoref{line:expand-recurse}).
%

@@

\section{Proof of characterization of CDD systems}
\label{app:char}
The following is a proof of \autoref{thm:cdd-contains}.
%
\begin{proof}
  % proof structure:
  To prove that CDD is a strict superset of the union of the class
  of linear systems and the class of body-disjoint systems, we prove
  %
  \textbf{(1)} CDD contains the class of linear systems,
  \textbf{(2)} CDD contains the class of body-disjoint systems, and
  \textbf{(3)} there is some CDD system that is neither linear nor
  body-disjoint.

  % contains linear systems:
  For goal \textbf{(1)}, let $S$ be an arbitrary linear
  system.
  %
  $S$ is CDD if for each clause $C$ in
  $S$, each pair of distinct predicates in
  the body of $C$, the transitive dependencies are disjoint and
  no predicates appears more than once in the body of $C$.
  (\autoref{defn:cdds}).
  %
  Let $C$ be an arbitrary clause in $S$.
  %
  Since $C$ is a linear clause, it has at most one relational
  predicate in its body. Therefore, the transitive dependencies are
  trivially disjoint and there can be no repeated predicate.
  %
  Therefore, $S$ is CDD.

  % contains body-disjoint systems:
  For goal \textbf{(2)}, let $S$ be an arbitrary
  body-disjoint system.
  %
  The dependence relation of $S$ is a tree $T$, by the
  definition of a body-disjoint system.
  %
  Let $C$ be an arbitrary clause in $S$, with
  distinct relational predicates $R_0$ and $R_1$ in its body.
  %
  All dependencies of $R_0$ and $R_1$ are in subtrees of $T$, which
  are disjoint by the definition of a tree.
  %
  Thus, $S$ is CDD, by \autoref{defn:cdds}.

  % distinguishing example:
  For goal \textbf{(3)}, the system $\mcchc$ is CDD, but is neither linear
  nor body-disjoint.
\end{proof}

\section{Proof $\solvecdd$ is in co-NP}
\label{app:solve-cp}
The following is a proof of \autoref{thm:solve-cp}.
%
Namely, the upper bound of $\solvecdd$ is in co-NP.
%
\begin{proof}
  $\prectr$ and $\postctr$ construct formulas linear in the size of the
  CHC system. The satisfiability problem for the constructed formulas
  are in NP for linear arithmetic.
  %
  $\solvecdd$ at issues (at worst) a linear number of interpolation
  queries in terms of number of predicate.
  %
  Therefore, the upper bound of $\solvecdd$ is co-NP.
\end{proof}

### 2018-04-13 paper/approach.tex
add the rough proof
pass over all except proof of correctness

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
CDD systems can be solved
in co-NP time and contains the union of all previous introduced classes
that can be solved in co-NP time.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all sibling pairs, $P, Q \in \predof{S}$,
  %
  the transitive dependences of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
  \tdepsof{Q} = \emptyset$)
  %
  and no predicate shows more than once in the body of a single clause,
  %
  then $\cc{S}$ is \emph{Clause-Dependence Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branches and procedure calls such
that each execution path invokes each statement at most once.
%
\begin{ex}
  The CHC system $\mcchc$ is a CDD system. An argument is given in
  \autoref{sec:solve-ex}.
\end{ex}

@@

% talk about linear system
\begin{thm}
\label{thm:cdd-contains}
  The class of CDD systems strictly contains the union of the classes
  of body-disjoint and linear systems.
\end{thm}
%
Proof is given in \autoref{app:char}.

@@

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd(\cc{S})$ %
    \label{line:solve-begin}}{ %
    $\sigma \assign \emptyset$ \\
    $\cc{Preds} \assign \topSort(\predof{S})$ \\
    \For {$P \in \cc{Preds}$}
    {$\cc{interpolant} \assign \solveitp(\prectr(P,\sigma),\postctr(P,\sigma))$\\
     \Switch{\cc{interpolant}}{
       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}
       \lCase{$\cc{I}$:}{$\sigma$ [$P$] $\assign$ $\cc{I}$}
     }
    }
    \Return{$\sigma$}
    }
  %
  \caption{$\solvecdd$: for a CDD system $\cc{S}$, returns a
    solution to $\cc{S}$ or the value $\none$ to denote that
    $\cc{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}

% introduce solveCDD
\autoref{alg:solve-cdd} presents $\solvecdd$, a procedure designed to solve
CDD systems.
%
Given a CDD system $\cc{S}$, $\solvecdd$ topologically sorts the uninterpreted
predicates in $\cc{S}$ based on their dependency relations.
%
Then, the algorithm calculates interpretations for each predicate in this order
by invoking $\solveitp$.
%
$\solveitp$ computes a binary interpolant of the pre and post formulas of the
given predicate, where these formulas are based on the current, partial
solution.
%
The pre and post formulas are computed respectively by $\prectr$ and
$\postctr$, which we define in \autoref{sec:cons-pre} and \autoref{sec:cons-post}.
%
It is possible that the pre and post formulas may be mutually satisifiable, in
which case $\solveitp$ returns \cc{SAT}. In this case, $\solvecdd$ returns
$\nosoln$ to indicate that $\cc{S}$ is not solvable.
%
Otherwise, $\solvecdd$ updates the partial solution by setting the
interpretation of $P$ to $\cc{I}$.
%
Once all predicates have been interpolated, $\solvecdd$ returns the complete
solution, $\sigma$.

@@

% solve-cdd in coNP
\begin{thm}
  \label{thm:solve-cp}
  Given a CDD system $S$, $\solvecdd$ either returns the solution of $S$ or $\none$
  in coNP time.
\end{thm}

\subsubsection{Constructing constraints for predicates}
%
For a CDD system $S$, a predicate $P \in \predof{S}$, and a partial
solution $\sigma$ that maps predicates to their solutions, the formula
$\ctrof{P,\sigma}$ is a compact representation of the constraints of
$P$.
%
If $\sigma$ does not contain $P$, then the constraint of $P$ is
constructed from the clauses where $P$ is the head.  When $\sigma$
does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
%
\[
  \ctrof{P,\sigma}=
  \begin{dcases}
    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = P)}
    \left( \consof{C_i} \land %
      \bigland_{ \cc{Q} \in \bodyof{C_i}} \cc{b}_Q
    \right),
  &\text{if } P \notin \sigma\\
  \sigma[P], &\text{if } P \in \sigma
  \end{dcases}
\]
%
% introduce the isused predicates:
Each $P \in \predof{S}$ has a corresponding boolean variable $\cc{b}_P$.
%
The counterexample characterization of $P$ constrains that if $P$ is used,
(meaning $\cc{b}_{P} = True$) then its constraint must hold:
\[
  \vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}
\]

\begin{ex}
  \label{ex:ctr}
  When $\solvecdd$ solves the predicate $\cc{L}_9$ in $\mcchc$, it generates a
  constraint based on clauses (5) and (6):
  $$\ctrof{\cc{L}_9,\sigma} =
    (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
    \lor
    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$$
  The counterexample characterization for $\cc{L}_9$ is based on its boolean
  indicator and this constraint:
  $$\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$$
\end{ex}

% pre-constraint:
\subsubsection{Constructing pre-formulas for predicates}
\label{sec:cons-pre}
%
$\prectr(P, \sigma)$ denotes the pre-formula for an arbitrary predicate $P$
with respect to the partial solution map, $\sigma$.
%
Due to the topological ordering, when $\solvecdd$ attempts to solve $P$, the
interpretations for all dependencies of $P$ will be stored in
$\sigma$.
%
The pre-formula is built from these interpretations together with boolean
indicators of the dependencies and the constraint of $P$:
%
\[
\prectr(P,\sigma) =
  \ctrof{P,\sigma} \land
    \left(
      \bigland_{\cc{Q} \in \depsof{P}}
      \left(\neg b_Q \lor \sigma[ \cc{Q} ]  \right)
    \right)
\]
%
\begin{ex}
  \label{ex:pre-ctr}
  When $\solvecdd$ solves the uninterpreted predicate $\cc{L}_9$ in $\mcchc$,
  $\sigma$ maps $\cc{L}_6$ to $n \ge 0$ and $\cc{L}_8$ to $n < 0$.
  %
  The pre-formula for $\cc{L}_9$ under $\sigma$ is therefore:
  $$\ctrof{\cc{L}_9}
  \land (\neg \cc{b}_{\cc{L}_6} \lor n \ge 0) \land (\neg \cc{b}_{\cc{L}_8}
  \lor n < 0)$$
  %
  The formula $\ctrof{\cc{L}_9,\sigma}$ is given in~\autoref{ex:ctr}.
\end{ex}

\subsubsection{Constructing post-formulas for predicates}
\label{sec:cons-post}
$\postctr(P, \sigma)$ denotes the post-formula for an arbitrary predicate $P$
with respect to the partial solution map, $\sigma$. A valid post-constraint is
mutually inconsistent with the solution of $P$.
%
The post formula is constructed based on the predicates which depend on $P$.
%
% dependents of R:
Let $D_0$ be the \emph{reflexive transitive dependents} of $P$ in $S$,
% clause siblings of dependents:
let $D_1$ be the siblings in $S$ of $(D_0 \union {P})$,
%
let $D_2$ be all transitive dependences of $D_1$, and
% collect all predicates that we define a constraint over:
let $D = D_0 \union D_1 \union D_2$.
% define instantiation of C
The post constraint for $P$ under $\sigma$ is the conjunction of counterexample
characterization of all predicates $Q \in D$ and the query clause.
%
\[
\postctr(R, \sigma) =\cc{query} \land (\bigland_{\cc{Q} \in D} \vc{Q,\sigma})
\]
% example post constraint:
\begin{ex}
  \label{ex:ctx-ctr}
  When $\solvecdd$ solves $\cc{L}_9$ in $\mcchc$, it must consider
  the dependents of $\cc{L}_9$.
  %
  The \emph{reflexive transitive dependents} $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
  %
  The siblings set $D_1$ is \{\cc{dbl}\}.
  %
  The transitive dependences of $D_1$ is $\varnothing$. 
  %
  Therefore, the $D$ set is \{\cc{main},\cc{dbl}\}.
  %
  The \cc{query} is $\cc{b}_{main} \land res<0$.
  %
  The post constraint for $\cc{L}_9$ under $\sigma$ is
  $$\cc{query} \land \vc{\cc{main},\sigma} \land \vc{\cc{dbl},\sigma}$$
\end{ex}

@@

%
\begin{defn}
  If $S'$ is CDD, $S \expandsto S'$, and there is no CDD system $S''$ such that $S \expandsto S' \expandsto S''$ and
  $S'' \neq S'$, then $S'$ is a minimal CDD expansion of $S$.
\end{defn}

>>>>

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
CDD systems can be solved
in co-NP time and contains the union of all previous introduced classes
that can be solved in co-NP time.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all sibling pairs, $P, Q \in \predof{S}$,
  %
  the transitive dependencies of $P$ and $Q$ are disjoint ($\tdepsof{P} \cap
  \tdepsof{Q} = \emptyset$)
  %
  and no predicate shows more than once in the body of a single clause,
  %
  then $\cc{S}$ is \emph{Clause-Dependence Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branches and procedure calls such
that each execution path invokes each statement at most once.
%
\begin{ex}
  The CHC system $\mcchc$ is a CDD system. An argument is given in
  \autoref{sec:solve-ex}.
\end{ex}

@@

% talk about linear system
\begin{thm}
\label{thm:cdd-contains}
  The class of CDD systems is a strict superset of the union of the
  class of body-disjoint systems and the class of linear systems.
\end{thm}
%
Proof is given in \autoref{app:char}.

@@

% algorithm for constructing a CDD system:
\begin{algorithm}[t]
  % Declare IO markers.
  \SetKwInOut{Input}{Input}
  %
  \SetKwInOut{Output}{Output}
  % Declare sub-program (procedure) markers.
  \SetKwProg{myproc}{Procedure}{}{}
  % Inputs: a heap program and an error location.
  \Input{A CDD System $\cc{S}$.}
  % Output: inductive invariants.
  \Output{If $\cc{S}$ is solvable, then a solution of
    $\cc{S}$; %
    otherwise, the value $\nosoln$.}
  % 
  \myproc{$\solvecdd(\cc{S})$ %
    \label{line:solve-begin}}{%
    $\sigma \assign \emptyset$ \\
    $\cc{Preds} \assign \topSort(\predof{S})$ \\
    \label{line:top-sort}}{%
    \For {$P \in \cc{Preds}$}
    {$\cc{interpolant} \assign \solveitp(\prectr(P,\sigma),\postctr(P,\sigma))$
      \label{line:interp}\\
     \Switch{\cc{interpolant}}{%
       \lCase{\cc{SAT}:}{\Return{$\nosoln$}}\label{line:interp-sat}
       \lCase{$\cc{I}$:}{$\sigma$ [$P$] $\assign$ $\cc{I}$}\label{line:interp-valid}
     }
    }
    \Return{$\sigma$}\label{line:solve-done}
    }
  %
  \caption{$\solvecdd$: for a CDD system $\cc{S}$, returns a
    solution to $\cc{S}$ or the value $\none$ to denote that
    $\cc{S}$ has no solution.}
  \label{alg:solve-cdd}
\end{algorithm}

% introduce solveCDD
\autoref{alg:solve-cdd} presents $\solvecdd$, a procedure designed to solve
CDD systems.
%
Given a CDD system $\cc{S}$, $\solvecdd$ topologically sorts the uninterpreted
predicates in $\cc{S}$ based on their dependency relations
(\autoref{line:top-sort}).
%
Then, the algorithm calculates interpretations for each predicate in this order
by invoking $\solveitp$ (\autoref{line:interp}).
%
$\solveitp$ computes a binary interpolant of the pre and post formulas of the
given predicate, where these formulas are based on the current, partial
solution.
%
The pre and post formulas are computed respectively by $\prectr$ and
$\postctr$, which we define in \autoref{sec:cons-pre} and \autoref{sec:cons-post}.
%
It is possible that the pre and post formulas may be mutually
satisifiable, in which case $\solveitp$ returns \cc{SAT}
(\autoref{line:interp-sat}). In this case, $\solvecdd$ returns
$\nosoln$ to indicate that $\cc{S}$ is not solvable.
%
Otherwise, $\solvecdd$ updates the partial solution by setting the
interpretation of $P$ to $\cc{I}$ (\autoref{line:interp-valid}).
%
Once all predicates have been interpolated, $\solvecdd$ returns the complete
solution, $\sigma$ (\autoref{line:solve-done}).

@@

% solve-cdd in coNP
\begin{thm}
  \label{thm:solve-cp}
  Given a CDD system $S$, $\solvecdd$ either returns the solution of $S$ or $\none$
  in co-NP time.
\end{thm}
Proof is given in \autoref{app:solve-cp}.

%
\subsubsection{Constructing constraints for predicates}
%
For a CDD system $S$, a predicate $P \in \predof{S}$, and a partial
solution $\sigma$ that maps predicates to their solutions, the formula
$\ctrof{P,\sigma}$ is a compact representation of the constraints of
$P$.
%
If $\sigma$ does not contain $P$, then the constraint of $P$ is
constructed from the clauses where $P$ is the head.  When $\sigma$
does contain $P$, $\ctrof{P,\sigma}$ is a lookup from $\sigma$:
%
\[
  \ctrof{P,\sigma}=
  \begin{dcases}
    \biglor_{(\cc{C}_i \in \cc{S}) \land (\headof{C_i} = P)}
    \left( \consof{C_i} \land %
      \bigland_{ \cc{Q} \in \bodyof{C_i}} \cc{b}_Q
    \right),
  &\text{if } P \notin \sigma\\
  \sigma[P], &\text{if } P \in \sigma
  \end{dcases}
\]
%
% introduce the isused predicates:
Each $P \in \predof{S}$ has a corresponding boolean variable $\cc{b}_P$.
%
The counterexample characterization of $P$ constrains that if $P$ is used
(meaning $\cc{b}_{P} = True$), then its constraint must hold:
\[
  \vc{P,\sigma} = \neg \cc{b}_{P} \lor \ctrof{P,\sigma}
\]

\begin{ex}
  \label{ex:ctr}
  When $\solvecdd$ solves predicate $\cc{L}_9$ in $\mcchc$, it generates a
  constraint based on clauses (5) and (6):
  $$\ctrof{\cc{L}_9,\sigma} =
    (\cc{abs'} = \cc{n} \land \cc{b}_{\cc{L}_6})
    \lor
    (\cc{abs'} = -\cc{n} \land \cc{b}_{\cc{L}_8})$$
  The counterexample characterization for $\cc{L}_9$ is based on its boolean
  indicator and this constraint:
  $$\vc{\cc{L}_9,\sigma} = \neg \cc{b}_{\cc{L}_9} \lor \ctrof{\cc{L}_9,\sigma}$$
\end{ex}

% pre-constraint:
\subsubsection{Constructing pre-formulas for predicates}
\label{sec:cons-pre}
%
$\prectr(P, \sigma)$ denotes the pre-formula for an arbitrary predicate $P$
with respect to the partial solution map, $\sigma$.
%
Due to the topological ordering, when $\solvecdd$ attempts to solve $P$, the
interpretations for all dependencies of $P$ will be stored in
$\sigma$.
%
The pre-formula is built from these interpretations together with boolean
indicators of the dependencies and the constraint of $P$:
%
\[
\prectr(P,\sigma) =
  \ctrof{P,\sigma} \land
    \left(
      \bigland_{\cc{Q} \in \depsof{P}}
      \left(\neg b_Q \lor \sigma[ \cc{Q} ]  \right)
    \right)
\]
%
\begin{ex}
  \label{ex:pre-ctr}
  When $\solvecdd$ solves predicate $\cc{L}_9$ in $\mcchc$,
  $\sigma$ maps $\cc{L}_6$ to $n \ge 0$ and $\cc{L}_8$ to $n < 0$.
  %
  The pre-formula for $\cc{L}_9$ under $\sigma$ is therefore:
  $$\ctrof{\cc{L}_9}
  \land (\neg \cc{b}_{\cc{L}_6} \lor n \ge 0) \land (\neg \cc{b}_{\cc{L}_8}
  \lor n < 0)$$
  %
  The formula $\ctrof{\cc{L}_9,\sigma}$ is given in~\autoref{ex:ctr}.
\end{ex}

\subsubsection{Constructing post-formulas for predicates}
\label{sec:cons-post}
$\postctr(P, \sigma)$ denotes the post-formula for an arbitrary predicate $P$
with respect to the partial solution map, $\sigma$. A valid
post-formula is mutually inconsistent with the solution of $P$.
%
and is constructed based on the predicates which depend on $P$.
%
% dependents of R:
Let $D_0$ be the \emph{reflexive transitive dependents} of $P$ in $S$,
% clause siblings of dependents:
let $D_1$ be the siblings in $S$ of $(D_0 \union {P})$,
%
let $D_2$ be all transitive dependencies of $D_1$, and
% collect all predicates that we define a constraint over:
let $D = D_0 \union D_1 \union D_2$.
% define instantiation of C
The post-formula for $P$ under $\sigma$ is the conjunction of counterexample
characterization of all predicates $Q \in D$ and the query clause:
%
\[
\postctr(R, \sigma) =\cc{query} \land (\bigland_{\cc{Q} \in D} \vc{Q,\sigma})
\]
% example post constraint:
\begin{ex}
  \label{ex:ctx-ctr}
  When $\solvecdd$ solves $\cc{L}_9$ in $\mcchc$, it must consider
  the dependents of $\cc{L}_9$.
  %
  The \emph{reflexive transitive dependents} $D_0$ of $\cc{L}_9$ is \{\cc{main}\}.
  %
  The siblings set $D_1$ is \{\cc{dbl}\}.
  %
  The transitive dependencies of $D_1$ is $\varnothing$. 
  %
  Therefore, the $D$ set is \{\cc{main},\cc{dbl}\}.
  %
  The \cc{query} is $\cc{b}_{main} \land res<0$.
  %
  The post-formula for $\cc{L}_9$ under $\sigma$ is:
  $$\cc{query} \land \vc{\cc{main},\sigma} \land \vc{\cc{dbl},\sigma}$$
\end{ex}

@@

%
\begin{defn}
  \label{defn:min-expansion}
  If $S'$ is CDD, $S \expandsto S'$, and there is no CDD system $S''$ such that $S \expandsto S' \expandsto S''$ and
  $S'' \neq S'$, then $S'$ is a minimal CDD expansion of $S$.
\end{defn}
